---
title: "GeneralPredictive2"
author: "Muhammad Athariq Naufal"
date: "6/9/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages()
```

```{r}
library(mlbench)
library(pROC)
library(rpart)

data("PimaIndiansDiabetes")
df <- PimaIndiansDiabetes
```


```{r}
BasicSummary <- function(df, dgts = 3){
m <- ncol(df)
varNames <- colnames(df)
varType <- vector("character",m)
topLevel <- vector("character",m)
topCount <- vector("numeric",m)
missCount <- vector("numeric",m)
levels <- vector("numeric", m)
for (i in 1:m){
x <- df[,i]
varType[i] <- class(x)
xtab <- table(x, useNA = "ifany")
levels[i] <- length(xtab)
nums <- as.numeric(xtab)
maxnum <- max(nums)
topCount[i] <- maxnum
maxIndex <- which.max(nums)
lvls <- names(xtab)
topLevel[i] <- lvls[maxIndex]
missIndex <- which((is.na(x)) | (x == "") | (x == " "))
missCount[i] <- length(missIndex)
}

n <- nrow(df)
topFrac <- round(topCount/n, digits = dgts)
missFrac <- round(missCount/n, digits = dgts)
summaryFrame <- data.frame(variable = varNames, type = varType,
levels = levels, topLevel = topLevel,
topCount = topCount, topFrac = topFrac,
missFreq = missCount, missFrac = missFrac)
return(summaryFrame)
}
```

```{r}
BasicSummary(df)
```

# levels
# toplevel
# topcount =  jumlah output toplevel (17 observasi yang hasilnya 99)
# topfrac = 
# missfreq = missing variable 
# missfrac = missing fraction dari missfreq

# Tricep and Insulin = tidak diikutsertakan 
# observing that the tricep and insulin are not qualified due to their th value are "0" (30% and 50%)
# binary observation is considered 

```{r}
keepvars <- c("pregnant","glucose", "pressure", "mass", "pedigree", "age", "diabetes")
PimaSub <- PimaIndiansDiabetes [, keepvars]
```


#1.   Replacing the respo

```{r}
set.seed(123) #menghasilkan bilangan acak

n <- nrow(PimaSub)
train <- sample(n, round(0.7 * n)) #70% dari row dfFiltered
pimaTrain <- PimaSub[train,] #mengambil 70% data secara acak
pimaValidation <- PimaSub[-train,] #mengambil sisa pimaTrain secara acak
```

#2.   Converting labels to discrete   

```{r}
diabetesFlag <- ifelse (PimaIndiansDiabetes$diabetes == 'neg', 0, 1)
PimaSub$diabetes <- diabetesFlag
```

#3.    Split data into training set and test set
```{r}
# Train a decision in tree model
DiabetesModel1 <- rpart(formula = diabetes ~.,
                        data = pimaTrain,
                        method = "class")
DiabetesModel1$variable.importance
```

```{r}
library(rpart.plot)
# yesno = 2, provides yes no 
rpart.plot (x = DiabetesModel1, yesno = 2, type = 0)
```


#4. Fit the model using full set of variables

```{r}
DiabetesModel2 <- rpart(formula = diabetes ~glucose + mass + age,
                        data = pimaTrain,
                        method = "class")
```
minsplit: mininmum number of observation that must exist in the node in order for a split to be attemppted
maxdepth: the maximum depth of any node of the final tree


#5. Model Validation
```{r}
PredModel1 <- predict(DiabetesModel1, newdata = pimaValidation, type = "class")
PredModel2 <- predict(DiabetesModel2, newdata = pimaValidation, type = "class")
```


#6. Model Performance
##6a. Confusion matrix for 1st model

```{r}
# Build confusion matrix w/ a threshold of 0.5

cm <- table(PredModel1, pimaValidation$diabetes)

accuracy <- sum (cm[1], cm[4]) / sum (cm[1:4])
precision <- cm[4] / sum (cm[4], cm[2])
sensitivity <- cm[4] / sum (cm[4], cm[3])
fscore <- (2 * (sensitivity * precision)) / (sensitivity * precision)
specificity <- cm[1] / sum(cm[1], cm[2])
sprintf("Accuracy  is %s", round(accuracy,3))
```

##6b. confusion matrix for second model
```{r}
cm <- table(PredModel2, pimaValidation$diabetes)
cm
accuracy <- sum (cm[1], cm[4]) / sum (cm[1:4])
precision <- cm[4] / sum (cm[4], cm[2])
sensitivity <- cm[4] / sum (cm[4], cm[3])
fscore <- (2 * (sensitivity * precision)) / (sensitivity + precision)
specificity <- cm[1] / sum(cm[1], cm[2])
sprintf("Accuracy  is %s", round(accuracy,3))
```

#7. conclusion







