---
title: "KP"
author: "Elvina"
date: "2023-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#("MLmetrics")

```

#1. load libraries
```{r}
library(mlbench)
library(pROC)
library(MLmetrics)
data("PimaIndiansDiabetes")

```

#2. 
```{r}
BasicSummary <- function(df, dgts = 3){
m <- ncol(df)
varNames <- colnames(df)
varType <- vector("character",m)
topLevel <- vector("character",m)
topCount <- vector("numeric",m)
missCount <- vector("numeric",m)
levels <- vector("numeric", m)
for (i in 1:m){
x <- df[,i]
varType[i] <- class(x)
xtab <- table(x, useNA = "ifany")
levels[i] <- length(xtab)
nums <- as.numeric(xtab)
maxnum <- max(nums)
topCount[i] <- maxnum
maxIndex <- which.max(nums)
lvls <- names(xtab)
topLevel[i] <- lvls[maxIndex]
missIndex <- which((is.na(x)) | (x == "") | (x == " "))
missCount[i] <- length(missIndex)
}
n <- nrow(df)
topFrac <- round(topCount/n, digits = dgts)
missFrac <- round(missCount/n, digits = dgts)
summaryFrame <- data.frame(variable = varNames, type = varType,
levels = levels, topLevel = topLevel,
topCount = topCount, topFrac = topFrac,
missFreq = missCount, missFrac = missFrac)
return(summaryFrame)
}

```


```{r}
BasicSummary(PimaIndiansDiabetes)

```
From the observation above, we can conclude that the TopLevel of Insulin and Triceps are "0" around 30-50% of the data .Therefore , the binary classifiers cosidered here omit both of these variables from the set.

```{r}
keepvar<-c("pregnant", "glucose", "pressure", "mass", "pedigree", "age", "diabetes")
pimasub <- PimaIndiansDiabetes[,keepvar]
BasicSummary(pimasub)
```

#3. Split data into training and test set
```{r}
set.seed(123)
n <- nrow(pimasub)
train <- sample(n, round(0.7 * n))
pimaTrain <- pimasub[train,]
pimaValidation<-pimasub[-train,]
```

```{r}
#install.packages("rpart")
#install.packages("rpart.plot")
```
#3. Model Building using full set variables
```{r}
#Train a decisiom tree model
library(rpart)
library(rpart.plot)
DiabetesModel1 <- rpart(formula = diabetes ~., data = pimaTrain, method = "class")
DiabetesModel1$variable.importance
```
tiap variabel yang kita sertakan di nilainya
semakain besar nilainya, semakin penting variabel tersebut

# Model Plotting
# Bentuk pohon keputusan
```{r}
rpart.plot(x = DiabetesModel1, yesno=2, type=0)
#yesno = 2 -> privides yes or no text
```

#4. Model that includes only 3 most important variables
# disini dipakai 3 teratas
```{r}
#train a decision model
DiabetesModel2 <- rpart(formula = diabetes ~ glucose + mass + age, data = pimaTrain, method = "class")
```
minsplit: the minimum number of observation that must exist in the node in order for a split to be attemped
maxdepth: the maximum depth of any node of the final tree

#5. Model validation
```{r}
PredModel1 <- predict(DiabetesModel1, newdata = pimaValidation, type = "class")
PredModel2 <- predict(DiabetesModel2, newdata = pimaValidation, type = "class")
```

#6. Model Performance
##6a. Confusion matrix for 1st model
```{r}
# Built confusion matrix with a threshold of 0.5
cm <- table(PredModel1, pimaValidation$diabetes)
cm
cm[2]

accuracy <- sum(cm[1], cm[4]) / sum(cm[1:4])
precision <- cm[4] / sum(cm[4], cm[2])
sensitivity <- cm[4] / sum(cm[4], cm[3])
fscore <- (2 * (sensitivity*precision)) / (sensitivity+precision)
specificity <- cm[1] / sum(cm[1], cm[2])
sprintf("Accuracy is %s", round(accuracy, 3))
```

##6b. Confusion matrix for 2nd model
```{r}
# Built confusion matrix with a threshold of 0.5
cm <- table(PredModel2, pimaValidation$diabetes)
cm
cm[2]

accuracy <- sum(cm[1], cm[4]) / sum(cm[1:4])
precision <- cm[4] / sum(cm[4], cm[2])
sensitivity <- cm[4] / sum(cm[4], cm[3])
fscore <- (2 * (sensitivity*precision)) / (sensitivity+precision)
specificity <- cm[1] / sum(cm[1], cm[2])
sprintf("Accuracy is %s", round(accuracy, 3))
```
The second model has higher accuracy than the first model.

#7. Conclusion
Since the second model involves only 3 prediction variables and provides higher accuracy than the first one. Moreover, it is substantially simpler model (i.e., it is easier to explain to others withl limited technical background).


