---
title: "Session 12_KP"
author: "Wilbert"
date: "2023-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(mlbench)
library(pROC)
library(MLmetrics)

data <- Boston
head(data)
```
```{r}
BasicSummary <- function(df, dgts = 3){
m <- ncol(df)
varNames <- colnames(df)
varType <- vector("character",m)
topLevel <- vector("character",m)
topCount <- vector("numeric",m)
missCount <- vector("numeric",m)
levels <- vector("numeric", m)
for (i in 1:m){
x <- df[,i]
varType[i] <- class(x)
xtab <- table(x, useNA = "ifany")
levels[i] <- length(xtab)
nums <- as.numeric(xtab)
maxnum <- max(nums)
topCount[i] <- maxnum
maxIndex <- which.max(nums)
lvls <- names(xtab)
topLevel[i] <- lvls[maxIndex]
missIndex <- which((is.na(x)) | (x == "") | (x == " "))
missCount[i] <- length(missIndex)
}

n <- nrow(df)
topFrac <- round(topCount/n, digits = dgts)
missFrac <- round(missCount/n, digits = dgts)
summaryFrame <- data.frame(variable = varNames, type = varType,
levels = levels, topLevel = topLevel,
topCount = topCount, topFrac = topFrac,
missFreq = missCount, missFrac = missFrac)
return(summaryFrame)
}
```

```{r}
BasicSummary(data)
summary(data)
```
From the observation above, we can conclude that the TopLevel of zn and chas are "0" around 30-50% of the data .Therefore , the binary classifiers cosidered here omit both of these variables from the set.

```{r}
keepvars <- c("crim", "indus", "nox", "rm", "age", "dis", "rad", "tax", "ptratio", "black", "lstat", "medv")
BostonSub <- Boston[, keepvars]
```

# Coverting Labels to Discrete
```{r}
medvFlag <- ifelse(Boston$medv <= 21, 0, 1)
BostonSub$medv <- as.factor(medvFlag)
```

# Separate Train and Test Data
```{r}
set.seed(123)

n <- nrow(BostonSub)
train <- sample(n, (0.7 * n)) # 70% of the data
bostonTrain <- BostonSub[train,]
bostonTest <- BostonSub[-train,]
```

# Train a Decission Tree
## Fitting the model with all of the variables
```{r}
medvModel1 <- rpart(formula = medv ~.,
                    data = bostonTrain,
                    method = "class")
medvModel1$variable.importance
```
```{r}
library(rpart.plot)
rpart.plot(x = medvModel1, yesno = 2, type = 0)
```

## Fit the model only with top 3 most important variables
```{r}
medvModel2 <- rpart(formula = medv ~ lstat + rm + nox,
                    data = bostonTrain,
                    method = "class")
medvModel2$variable.importance
```

# Model Validation
```{r}
predModel1 <- predict(medvModel1, newdata = bostonTest, type = "class")
predModel2 <- predict(medvModel2, newdata = bostonTest, type = "class")
```

# Model Performance
```{r}
confusionMatrix <- function(Model, testData) {
  cm <- table(Model, testData)

  accuracy <- sum (cm[1], cm[4]) / sum (cm[1:4])
  precision <- cm[4] / sum (cm[4], cm[2])
  sensitivity <- cm[4] / sum (cm[4], cm[3])
  fscore <- (2 * (sensitivity * precision)) / (sensitivity * precision)
  specificity <- cm[1] / sum(cm[1], cm[2])
  sprintf("Accuracy  is %s", round(accuracy,3))
}
```

## Confusion Matrix for 1st Model
```{r}
confusionMatrix(predModel1, bostonTest$medv)
```

## Confusion Matrix for 2nd Model
```{r}
confusionMatrix(predModel2, bostonTest$medv)
```

## Conclusion
Since the second model involves only 3 prediction variables and provides higher accuracy than the first one. Moreover, it is substantially simpler model (i.e., it is easier to explain to others withl limited technical background).











