---
title: "Regression Model"
author: "Wilbert"
date: "2023-06-14"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Regression vs Classification

- Regression -> Continuous
- Classification -> Discrete (Labels & Categorical)

Regression and Classification is differentiated by it's dependent variable.

# 1. Data Description

https://archive.ics.uci.edu/dataset/9/auto+mpg

    1. mpg:           continuous (Miles per Gallon, in gallon)
    2. cylinders:     multi-valued discrete (Number of Cylinders)
    3. displacement:  continuous (Engine Capacity)
    4. horsepower:    continuous (Engine Power)
    5. weight:        continuous (Vehicle Weight)
    6. acceleration:  continuous (Vehicle Acceleration)
    7. model year:    multi-valued discrete
    8. origin:        multi-valued discrete
    9. car name:      string (unique for each instance)
    
# 2. Load libraries & data
```{r}
library(rpart)
library(rpart.plot)
library(MLmetrics) 
# Regression use RMSF & Rsquared, pROC -> ROC
# Classification Sensivity, Specificity, Accuracy

autompg <- read.table("auto-mpg.data")
colnames(autompg) <- c("mpg", "cylinders", "displacement", "horsepower", "weight", "acceleration", "model_year", "origin", "car_name")
head(autompg)
```

# 3. Statistics Summary
```{r}
basicSummary <- function(df, dgts = 3) {
  m <- ncol(df) # column
  varNames <- colnames(df)
  varType <- vector("character", m)
  topLevel <- vector("character", m)
  topCount <- vector("numeric", m)
  missCount <- vector("numeric", m)
  levels <- vector("numeric", m)
  
  for (i in 1:m) {
    x <- df[,i]
    varType[i] <- class(x)
    xtab <- table(x, useNA = "ifany")
    levels[i] <- length(xtab)
    nums <- as.numeric(xtab)
    maxnum <- max(nums)
    topCount[i] <- maxnum
    maxIndex <- which.max(nums)
    lvls <- names(xtab)
    topLevel[i] <- lvls[maxIndex]
    missIndex <- which((is.na(x)) | (x == "") | (x == " "))
    missCount[i] <- length(missIndex)
  }
  n <- nrow(df) # Rows
  topFrac <- round(topCount/n, digits = dgts)
  missFrac <- round(missCount/n, digits = dgts)
  summaryFrame <- data.frame(variable = varNames, 
                             type = varType, 
                             levels = levels,
                             topLevel = topLevel,
                             topCount = topCount,
                             topFrac = topFrac,
                             missFreq = missCount,
                             missFrac = missFrac)
  return(summaryFrame)
}
```
```{r}
basicSummary(autompg)
# Because Horse wpoer is not the right data type, we must change it to the correct data type
```
# 4. Data Preparation

In Summary : 
Data Problems.

1. car_name is meaningless -> "remove"
2. convert all var into numerical
3. Horsepower have missng value

Rule of thumb : if missing value < 20% of total data then we can remove it, if not then median is recomended
```{r}
# Remove car_name
autompg <- subset(autompg, select = -c(car_name))

# Change all var into numerical
autompg$horsepower <- as.numeric(autompg$horsepower)
autompg$cylinders <- as.numeric(autompg$cylinders)
autompg$model_year <- as.numeric(autompg$model_year)
autompg$origin <- as.numeric(autompg$origin)

# Change missing value with mean
autompg$horsepower[is.na(autompg$horsepower)] <- mean(autompg$horsepower, na.rm = T)
basicSummary(autompg)
```
Warning -> becasue some of the character strings aren't properly structures integrers and so can't be translated to numeric


# 5. Split data 
```{r}
set.seed(123)
n <- nrow(autompg)
train <- sample(n, round(0.7 * n))
autoTrain <- autompg[train,]
autoTest <- autompg[-train,]
```

# 6. Fit the Model
## 6A. Linear Regression Models
```{r}
LRModel1 <- lm(mpg ~., autoTrain) # Model 1
#summary(LRModel1)

LRModel2 <- lm(mpg ~ weight + model_year + origin, autoTrain) # Model2
```

## 6B. Decission Tree Models
```{r}
DTModel1 <- rpart(mpg ~ ., autoTrain) # Model 1
#DTModel1$variable.importance
DTModel2 <- rpart(mpg ~ displacement + weight + cylinders + horsepower, autoTrain) # Model 2
```
Getting top 4 for the 2nd Model
```{r}
rpart.plot(DTModel1)
rpart.plot(DTModel2)
```

# 7. Model Validation
```{r}
predLR1 <- predict(LRModel1, autoTest)
predLR2 <- predict(LRModel2, autoTest)
predDT1 <- predict(DTModel1, autoTest)
predDT2 <- predict(DTModel2, autoTest)
```

# 8. Model Performance
```{r}
# Calculating RMSE
valLR1 <- RMSE(predLR1, autoTest$mpg)
valLR2 <- RMSE(predLR2, autoTest$mpg)
valDT1 <- RMSE(predDT1, autoTest$mpg)
valDT2 <- RMSE(predDT2, autoTest$mpg)

sprintf("RMSE of Linear Model 1 is %s", valLR1)
sprintf("RMSE of Linear Model 2 is %s", valLR2)
sprintf("RMSE of Decission Tree Model 1 is %s", valDT1)
sprintf("RMSE of Decission Tree Model 2 is %s", valDT2)

# Calculationg R2_Score
valLR1 <- R2_Score(predLR1, autoTest$mpg)
valLR2 <- R2_Score(predLR2, autoTest$mpg)
valDT1 <- R2_Score(predDT1, autoTest$mpg)
valDT2 <- R2_Score(predDT2, autoTest$mpg)

sprintf("R2_Score of Linear Model 1 is %s", valLR1)
sprintf("R2_Score of Linear Model 2 is %s", valLR2)
sprintf("R2_Score of Decission Tree Model 1 is %s", valDT1)
sprintf("R2_Score of Decission Tree Model 2 is %s", valDT2)
```

# 9. Conclusion
We can conclude that the best model is the Linear Regression Model 2 because it has the least RMSE and the highest R2_Score



