---
title: "2602113666_LabAssignment1"
author: "Wilbert Chandra"
date: "March 29, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Data
```{r}
library(ggplot2)
data("msleep")

msleep
```

User Defined Function to give basic summary
```{r}
basicSummary <- function(df, dgts) {
  m <- ncol(df) # column
  varNames <- colnames(df)
  varType <- vector("character", m)
  topLevel <- vector("character", m)
  topCount <- vector("numeric", m)
  missCount <- vector("numeric", m)
  levels <- vector("numeric", m)
  
  for (i in 1:m) {
    x <- df[,i]
    varType[i] <- class(x)
    xtab <- table(x, useNA = "ifany")
    levels[i] <- length(xtab)
    nums <- as.numeric(xtab)
    maxnum <- max(nums)
    topCount[i] <- maxnum
    maxIndex <- which.max(nums)
    lvls <- names(xtab)
    topLevel[i] <- lvls[maxIndex]
    missIndex <- which((is.na(x)) | (x == "") | (x == " "))
    missCount[i] <- length(missIndex)
  }
  n <- nrow(df) # Rows
  topFrac <- round(topCount/n, digits = dgts)
  missFrac <- round(missCount/n, digits = dgts)
  summaryFrame <- data.frame(variable = varNames, 
                             type = varType, 
                             levels = levels,
                             topLevel = topLevel,
                             topCount = topCount,
                             topFrac = topFrac,
                             missFraq = missCount,
                             missFrac = missFrac)
  return(summaryFrame)
}
```

# General Statistics
```{r}
str(msleep)
basicSummary(msleep)
table(msleep$conservation)
table(msleep$sleep_rem)
table(msleep$sleep_cycle)
table(msleep$brainwt)
```

From the table and descrition above we can see that the msleep dataset have 83 Objects and 11 Variables which is

- name : Categorical data with 83 unique values with each unique values having occuring 1 times. With no missing values
- genus : Categorical data with 77 unique values with "Pantera" being the most occuring value, occuring 3 times. With no missing values.
- vore : Categorical data with 5 unique values with "herbi" being the most occuring value, occuring 32 times. With no missing values.
- order : Categorical data with 19 unique values with "Rodentia" being the most occuring value, occuring 22 times. With no missing values.
- conservartion : Categorical data with 7 unique values with "lc" being the most occuring value ignoring missing value, occuring 27 times. With 29 missing values.
- sleep_total : Numerical data with 65 unique values with "12.5" being the most occuring values, occuring 4 times. With no missing values.
- sleep_rem : Numerical data with 33 unique values with "1.4" being the most occuring value ignoring missing value, occuring 5 times. With 22 missing values.
- sleep_cycle : Numerical data with 23 unique values with "0.183333333" being the most occuring value ignoring missing value, occuring 3 times. with 51 missing values.
- awake : Numerical data with 65 unique values with "11.5" being the most occuring value, occuring 4 times. with no missing values.
- brainwt : Numerical data with 54 unique values with "0.001", "0.115", and "0.18" being equally the most occuring values ignoring missing values, each occuring 2 times. with 27 missing values.
- bodywt : Numerical data with 82 unique values with "1.1" being the most occuring value, occuring 2 times. With no missing values.

# Descriptive Statistics
```{r}
summary(msleep)
```
The Descriptive Statistics of the msleep Datasets :

- sleep_total : with minimum of 1.90, median of 10.10, mean of 10.43, max of 19.90, 1st Quartile of 7.85, 3rd Quaritle of 13.75, and IQR of 5.9
- sleep_rem : with minimum of 0.1, median of 1.5, mean of 1.875, max of 6.6, 1st Quartile of 0.9, 3rd Quaritle of 2.4, and IQR of 1.5. And 22 missing values.
- sleep_cycle : with minimum of 0.1167, median of 0.3333, mean of 0.4396, max of 1.5, 1st Quartile of 0.1833, 3rd Quaritle of 0.5792, and IQR of 0.3859. And 51 missing values.
- awake : with minimum of 4.10, median of 13.90, mean of 13.57, max of 22.10, 1st Quartile of 10.25, 3rd Quaritle of 16.15, and IQR of 5.9
- brainwt : with minimum of 0.00014, median of 0.01240, mean of 0.28158, max of 5.712, 1st Quartile of 0.00290, 3rd Quaritle of 0.12550, and IQR of 0.1226. And 27 missing values.
- bodywt : with minimum of 0.005, median of 1.670, mean of 166.136, max of 6654.00, 1st Quartile of 0.174, 3rd Quaritle of 41.750, and IQR of 41.576

# Visualization
```{r}
plot(msleep, las = 1)
```
From the plot above I decided to explain more on the corelation between bodywt with vore, brainwt with vore, sleep_total with awake, and bodywt with bodywt. 

```{r}
ggplot(msleep, aes(x = bodywt, y = brainwt)) + geom_point() + labs(title = "Body Weight vs Brain Weight", x = "Body Weight", y = "Brain Weight") + scale_x_log10() + scale_y_log10()
```
From the Plot above we can see that there's a postive realtionn between body weight and brain weight thus we can conlude that the heavier an animal is then its brain will also be heavier.

```{r}
ggplot(msleep, aes(x = sleep_total, y = awake)) + geom_point() + labs(title = "Sleep Total vs Awake Time", x = "Sleep Total", y = "Awake Time")
```
From the plot above we can see that Sleep Total and Awake time have a negative relatiosip. Meaning that the more an animal sleep then it will be awake lesser and vice versa.

```{r}
ggplot(msleep, aes(x = vore, y = bodywt)) + geom_bar(stat = "summary", fun = "mean") + labs(title = "Body Weight Comparison with Vore Type", x = "Vore", y = "Body Weight" )
```
From the Graph above we can see that on average Herbivore have the heaviest body weight with Carnivore following. With Insectivore and Omnivore being the equally the lightest weight out of all.

```{r}
ggplot(msleep, aes(x = vore, y = brainwt)) + geom_bar(stat = "summary", fun = "mean") + labs(title = "Brain Weight Comparison with Vore Type", x = "Vore", y = "Brain Weight" )
```
From the graph above we can also see that on average Herbivore having the heaviest brain out of all. WIth Omnivore folloing, having the second heaviest brain on average and Carnivore being the 3rd heaviest brain and insectivore having the lightest brain out of all on average. 

# Anomalies
User Defined Function to Calculate How many anomalies
```{r}
extreme_values <- function(data) {
  
  # Calculate the IQR of the variable
  iqr <- IQR(data, na.rm = TRUE)
  
  # Calculate the lower and upper bounds for outliers
  lower_bound <- quantile(data, 0.25, na.rm = TRUE) - 1.5 * iqr
  upper_bound <- quantile(data, 0.75, na.rm = TRUE) + 1.5 * iqr
  
  # Count the number of extreme values
  extreme <- data[data < lower_bound | data > upper_bound]

  return(extreme)
}
```

```{r}
extreme_values(msleep$sleep_total)
```
There's no Extreme Values in sleep_total Variable

```{r}
extreme_values(msleep$sleep_cycle)
```
The sleep_cycle Variable has 2 extreme values of 1.5 and 1.416667

```{r}
extreme_values(msleep$sleep_rem)
```
The sleep_rem Variable has 3 extreme values of 4.9, 6.1, and 6.6.

```{r}
extreme_values(msleep$awake)
```
The awake Variable has no extreme value.

```{r}
extreme_values(msleep$brainwt)
```
The brainwt Variable has 8 extreme values of 0.423, 4.603, and 0.655, 0.419, 0.325, 1.320, 5.712, 0.440.

```{r}
extreme_values(msleep$bodywt)
```
The bodywt Variable has 11 extreme values of 600.000, 2547.000, 521.000,  187.000,  899.995,  800.000, 6654.000, 162.564, 161.499, 207.501,  173.330.