---
title: "Project-WB-Personal-Bayes"
author: "Wilbert"
date: "2023-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load rjags
```{r}
library(coda)
library(rjags)
```

# Load Function
```{r}
SVSS_cal <- function(beta) {
  library(knitr)
  Inc_Prob <- apply(beta!=0,2,mean)
  Q <- t(apply(beta,2,quantile,c(0.5,0.05,0.95)))
  out <- cbind(Inc_Prob,Q)
  kable(round(out,2))
}
```

```{r}
DIC_WAIC_cal <- function(model) {
  # Compute DIC
  dic1 <- dic.samples(model,n.iter=n.iter)
  # Compute WAIC
  waic1 <- coda.samples(model, variable.names=c("like"), n.iter=n.iter)
  like1 <- waic1[[1]]
  fbar1 <- colMeans(like1)
  P1 <- sum(apply(log(like1),2,var))
  WAIC1 <- -2*sum(log(fbar1))+2*P1
  
  print(dic1)
  print(WAIC1)
}
```

# Load Data
```{r}
df <- read.csv('./ParisHousing.csv')
head(df)
```

## Set Dataframe as matrix and split dependent and independent variable 
```{r}
price <- as.matrix(df$price)
Y <- price
X <- cbind(df$squareMeters, df$numberOfRooms, df$hasYard, df$hasPool, df$floors, df$cityCode, df$numPrevOwners, df$made, df$isNewBuilt, df$hasStormProtector, df$basement, df$attic, df$garage, df$hasStorageRoom, df$hasGuestRoom, df$cityPartRange)
names <- c("price","Square Meters", "Number of Rooms", "Has Yard", "Has Pool", "Number of Floors", "City Code", "Number of Previous Owners", "Year Made", "is Newly Built", "Has Storm Protector", "Basement Area", "Attic Area", "Garage Area", "Has Storage Room", "Has Guest Room", "City Part Range")

```

## Delete Missing Value
```{r}
junk <- is.na(rowSums(X))
Y <- Y[!junk]
X <- X[!junk,]
```

## Plot Data
```{r}
library(purrr)
library(tidyr)
library(ggplot2)

df %>%
  keep(is.numeric) %>%
  gather() %>%
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram()
```

## Standardize Covariates
```{r}
X <- as.matrix(scale(X))
```


# JAGS
## Put Data in JAGS Format
```{r}
n <- length(Y)
p <- ncol(X)

data <- list(Y=Y,X=X,n=n,p=p)
params <- c("alpha","beta")
burn <- 100
n.iter <- 2000
n.chains <- 3
```

## Make Jags Model
```{r}
model_string <- textConnection("model{
# Likelihood
for(i in 1:n){
Y[i] ~ dnorm(alpha+mu[i],taue)
mu[i] <- inprod(X[i,],beta[])
}
# Priors
for(j in 1:p){
beta[j] ~ dnorm(0,0.001)
}
alpha ~ dnorm(0,0.001)
taue ~ dgamma(0.1, 0.1)

# WAIC calculations
for(i in 1:n){
like[i] <- dnorm(Y[i],mu[i],taue)
}

}")
```

### Set Initial Value
```{r}
inits = list()

inits$alpha = rnorm(1)
for(i in 1:p) {
    inits$beta[i] = rnorm(1)
}
inits$taue = 10
```

### Compile Model
```{r}
model <- jags.model(model_string,data = data, n.chains=n.chains,quiet=TRUE, inits = inits)
```

### Update Model
```{r}
update(model, burn)
```

### Get Posterior Samples from the model
```{r}
samples1 <- coda.samples(model, variable.names=params, n.iter=n.iter)
```
### Plot MCMC Chain Trace and Features Posterior Density
```{r}
par(mar=c(1,1,1,1))
plot(samples1)
```
## Check Convergence
```{r}
gelman.diag(samples1)
```
The Chain in our model is convergent based on Gelma-Rubin Method with a value of 1.01

### Descriptive Statistics of model1
```{r}
summary(samples1)
```

### Show Alpha and Betas to the corresponding Features
```{r}
sum <- summary(samples1)
rownames(sum$statistics) <- names
rownames(sum$quantiles) <- names
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

### Autocorr
```{r}
autocorr.plot(samples1)
```


## Selection

```{r}
# Compile results
ESS1 <- effectiveSize(samples1)
out1 <- summary(samples1)$quantiles
rownames(out1)<-names

ESS1
```


### Plot Posterior Density Individually
```{r}
beta <- NULL
for(l in 1:n.chains){
beta <- rbind(beta,samples1[[l]])
}
colnames(beta) <- names
for(j in 1:17){
hist(beta[,j],xlab=expression(beta[j]),ylab="Posterior density",
breaks=100,main=names[j])
}
```

### SSVS
```{r}
library(knitr)
Inc_Prob <- apply(beta!=0,2,mean)
Q <- t(apply(beta,2,quantile,c(0.5,0.05,0.95)))
out <- cbind(Inc_Prob,Q)
kable(round(out,2))
```
Summary: All are included with posterior probability one and are
thus clearly important predictors of price. SO we decided to use all features

### DIC-WAIC
```{r}
# Compute DIC
dic1 <- dic.samples(model,n.iter=n.iter)
# Compute WAIC
waic1 <- coda.samples(model, variable.names=c("like"), n.iter=n.iter)
like1 <- waic1[[1]]
fbar1 <- colMeans(like1)
P1 <- sum(apply(log(like1),2,var))
WAIC1 <- -2*sum(log(fbar1))+2*P1
```

```{r}
dic1
WAIC1
```

# Model 2 -> With only Correlated Features

## Get Correlation Heatmap
```{r}
corr_mat <- round(cor(df),2) 

# Melted Corr Mat
library(reshape2)
melted_corr_mat <- melt(corr_mat)

# plotting the correlation heatmap
library(ggplot2)
ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2,
                                   fill=value)) + 
geom_tile()
```
From the Heatmap above we can see that price is only correlated with squareMeters

## Make into List
## split dependent and independent variable 
```{r}
X <- df$squareMeters
Y <- df$price
names <- c("price","Square Meters")

```

## Delete Missing Value
```{r}
junk <- is.na(X)
X <- X[!junk]
```


## JAGS
### Put Data in JAGS Format
```{r}
n <- length(Y)
n.chains <- 3
burn <- 500
n.iter <- 2000

data <- list(Y=Y,X=X,n=n)
params <- c("alpha","beta")
```

### Set Initial Value
```{r}
inits = list()

inits$alpha = rnorm(1)
inits$beta = rnorm(1)
inits$taue = 10
```

## Make Jags Model
```{r}
model_string <- textConnection("model{
# Likelihood
for(i in 1:n){
Y[i] ~ dnorm(alpha + beta*X[i],taue)
}

# Priors
beta ~ dnorm(0,0.001)
alpha ~ dnorm(0,0.001)
taue ~ dgamma(0.1, 0.1)

# WAIC calculations
for(i in 1:n){
like[i] <- dnorm(Y[i],beta * X[i],taue)
}

}")
```

### Compile Model
```{r}
model2 <- jags.model(model_string,data = data, n.chains=n.chains,quiet=TRUE, inits = inits)
```

### Update Model
```{r}
update(model2, burn)
```

### Get Posterior Samples from the model
```{r}
samples2 <- coda.samples(model2, variable.names=params, n.iter=n.iter)
```
### Plot MCMC Chain Trace and Features Posterior Density
```{r}
par(mar=c(1,1,1,1))
plot(samples2)
```
## Check Convergence
```{r}
gelman.diag(samples2)
```
The Chain in our model is convergent based on Gelma-Rubin Method with a value of 1.01

### Descriptive Statistics of model1
```{r}
summary(samples2)
```

### Show Alpha and Betas to the corresponding Features
```{r}
sum <- summary(samples2)
rownames(sum$statistics) <- names
rownames(sum$quantiles) <- names
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

### Autocorr
```{r}
autocorr.plot(samples2)
```


## Selection

```{r}
# Compile results
ESS2 <- effectiveSize(samples2)
out2 <- summary(samples2)$quantiles
rownames(out2)<-names

ESS2
```


### Plot Posterior Density Individually
```{r}
beta <- NULL
for(l in 1:n.chains){
beta <- rbind(beta,samples2[[l]])
}
colnames(beta) <- names
for(j in 1:2){
hist(beta[,j],xlab=expression(beta[j]),ylab="Posterior density",
breaks=100,main=names[j])
}
```
### SVSS
```{r}
SVSS_cal(beta)
```
Summary: All are included with posterior probability one and are
thus clearly important predictors of price. SO we decided to use all features

### DIC-WAIC

```{r}
DIC_WAIC_cal(model2)
```


# Frequentist Approach - Using only squareMeters
```{r}
head(df)
```

```{r}
plot(price ~ squareMeters, data = df)
```
```{r}
price.lm <- lm(price ~ squareMeters, data = df)

summary(price.lm)
```

# Model3 -> Initial Value from Frequentist Estimate
## Set Dataframe as matrix and split dependent and independent variable 
```{r}
X <- df$squareMeters
Y <- df$price
names <- c("price","Square Meters")

```

## Delete Missing Value
```{r}
junk <- is.na(X)
X <- X[!junk]
```


## JAGS
### Put Data in JAGS Format
```{r}
n <- length(Y)
n.chains <- 3
burn <- 500
n.iter <- 2000

data <- list(Y=Y,X=X,n=n)
params <- c("alpha","beta")
```

### Set Initial Value
```{r}
inits = list()

inits$alpha = 6461
inits$beta = 100
inits$taue = 10
```

## Make Jags Model
```{r}
model_string <- textConnection("model{
# Likelihood
for(i in 1:n){
Y[i] ~ dnorm(alpha + beta*X[i],taue)
}

# Priors
beta ~ dnorm(0,0.001)
alpha ~ dnorm(0,0.001)
taue ~ dgamma(0.1, 0.1)

# WAIC calculations
for(i in 1:n){
like[i] <- dnorm(Y[i],beta * X[i],taue)
}

}")
```

### Compile Model
```{r}
model3 <- jags.model(model_string,data = data, n.chains=n.chains,quiet=TRUE, inits = inits)
```

### Update Model
```{r}
update(model3, burn)
```

### Get Posterior Samples from the model
```{r}
samples3 <- coda.samples(model3, variable.names=params, n.iter=n.iter)
```
### Plot MCMC Chain Trace and Features Posterior Density
```{r}
par(mar=c(1,1,1,1))
plot(samples3)
```
## Check Convergence
```{r}
gelman.diag(samples3)
```
The Chain in our model is convergent based on Gelma-Rubin Method with a value of 1.01

### Descriptive Statistics of model1
```{r}
summary(samples3)
```

### Show Alpha and Betas to the corresponding Features
```{r}
sum <- summary(samples3)
rownames(sum$statistics) <- names
rownames(sum$quantiles) <- names
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

### Autocorr
```{r}
autocorr.plot(samples3)
```


## Selection

```{r}
# Compile results
ESS3 <- effectiveSize(samples3)
out3 <- summary(samples3)$quantiles
rownames(out3)<-names

ESS3
```


### Plot Posterior Density Individually
```{r}
beta <- NULL
for(l in 1:n.chains){
beta <- rbind(beta,samples3[[l]])
}
colnames(beta) <- names
for(j in 1:2){
hist(beta[,j],xlab=expression(beta[j]),ylab="Posterior density",
breaks=100,main=names[j])
}
```
### SVSS
```{r}
SVSS_cal(beta)
```
Summary: All are included with posterior probability one and are
thus clearly important predictors of price. SO we decided to use all features

### DIC-WAIC

```{r}
DIC_WAIC_cal(model3)
```

# Model 4 (Model 3 but with thinning -> 5)
## Set Dataframe as matrix and split dependent and independent variable 
```{r}
X <- df$squareMeters
Y <- df$price
names <- c("price","Square Meters")

```

## Delete Missing Value
```{r}
junk <- is.na(X)
X <- X[!junk]
```


## JAGS
### Put Data in JAGS Format
```{r}
n <- length(Y)
n.chains <- 3
burn <- 500
n.iter <- 2000
thin <- 5

data <- list(Y=Y,X=X,n=n)
params <- c("alpha","beta")
```

### Set Initial Value
```{r}
inits = list()

inits$alpha = 6461
inits$beta = 100
inits$taue = 10
```

## Make Jags Model
```{r}
model_string <- textConnection("model{
# Likelihood
for(i in 1:n){
Y[i] ~ dnorm(alpha + beta*X[i],taue)
}

# Priors
beta ~ dnorm(0,0.001)
alpha ~ dnorm(0,0.001)
taue ~ dgamma(0.1, 0.1)

# WAIC calculations
for(i in 1:n){
like[i] <- dnorm(Y[i],beta * X[i],taue)
}

}")
```

### Compile Model
```{r}
model4 <- jags.model(model_string,data = data, n.chains=n.chains,quiet=TRUE, inits = inits)
```

### Update Model
```{r}
update(model4, burn)
```

### Get Posterior Samples from the model
```{r}
samples4 <- coda.samples(model4, variable.names=params, n.iter=n.iter, thin = thin)
```
### Plot MCMC Chain Trace and Features Posterior Density
```{r}
par(mar=c(1,1,1,1))
plot(samples4)
```
## Check Convergence
```{r}
gelman.diag(samples4)
```
The Chain in our model is convergent based on Gelma-Rubin Method with a value of 1.01

### Descriptive Statistics of model1
```{r}
summary(samples4)
```

### Show Alpha and Betas to the corresponding Features
```{r}
sum <- summary(samples4)
rownames(sum$statistics) <- names
rownames(sum$quantiles) <- names
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

### Autocorr
```{r}
autocorr.plot(samples4)
```


## Selection

```{r}
# Compile results
ESS3 <- effectiveSize(samples4)
out3 <- summary(samples4)$quantiles
rownames(out3)<-names

ESS3
```


### Plot Posterior Density Individually
```{r}
beta <- NULL
for(l in 1:n.chains){
beta <- rbind(beta,samples4[[l]])
}
colnames(beta) <- names
for(j in 1:2){
hist(beta[,j],xlab=expression(beta[j]),ylab="Posterior density",
breaks=100,main=names[j])
}
```
### SVSS
```{r}
SVSS_cal(beta)
```
Summary: All are included with posterior probability one and are
thus clearly important predictors of price. SO we decided to use all features

### DIC-WAIC

```{r}
DIC_WAIC_cal(model4)
```

# Model 5 -> With only Correlated Features (3 stars)
```{r}
price.lm <- lm(price ~., data = df)

summary(price.lm)
```
Terdapat 7 variabel yang memiliki korelasi kuat dengan variabel price (dengan 3 stars). Ketujuh variabel tersebut antara lain squareMeters, hasYard, hasPool, floors, cityPartRange, isNewBuilt, dan hasStormProtector.

## Make into List
## split dependent and independent variable 
```{r}
X <- df[, c("squareMeters", "hasYard", "hasPool", "floors", "cityPartRange", "isNewBuilt", "hasStormProtector")]
Y <- df$price
names <- c("price", "Square Meters", "Has Yard", "Has Pool", "Floors", "City Part Range", "Is New Built", "Has Storm Protector")
```

# JAGS
## Put Data in JAGS Format
```{r}
n <- length(Y)
p <- ncol(X)

data <- list(Y=Y,X=X,n=n,p=p)
params <- c("alpha","beta")
burn <- 100
n.iter <- 2000
n.chains <- 3
```

## Make Jags Model
```{r}
model_string <- textConnection("model{
  # Likelihood
  for(i in 1:n){
    Y[i] ~ dnorm(alpha+mu[i],taue)
    mu[i] <- inprod(X[i,],beta[])
  }
  
  # Priors
  for(j in 1:p){
    beta[j] ~ dnorm(0,0.001)
  }
  
  alpha ~ dnorm(0,0.001)
  taue ~ dgamma(0.1, 0.1)
  
  # WAIC calculations
  for(i in 1:n){
    like[i] <- dnorm(Y[i],mu[i],taue)
  }

}")
```

### Set Initial Value
```{r}
inits = list()

inits$alpha = rnorm(1)
for(i in 1:p) {
    inits$beta[i] = rnorm(1)
}
inits$taue = 10
```

### Compile Model
```{r}
model5 <- jags.model(model_string,data = data, n.chains=n.chains,quiet=TRUE, inits = inits)
```

### Update Model
```{r}
update(model5, burn)
```

### Get Posterior Samples from the model
```{r}
samples5 <- coda.samples(model5, variable.names=params, n.iter=n.iter)
```
### Plot MCMC Chain Trace and Features Posterior Density
```{r}
par(mar=c(1,1,1,1))
plot(samples5)
```
## Check Convergence
```{r}
gelman.diag(samples5)
```
The Chain in our model is convergent based on Gelman-Rubin Method with a value of 1


### Descriptive Statistics of model1
```{r}
summary(samples5)
```

### Show Alpha and Betas to the corresponding Features
```{r}
sum <- summary(samples5)
rownames(sum$statistics) <- names
rownames(sum$quantiles) <- names
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

### Autocorr
```{r}
autocorr.plot(samples5)
```
## Selection

```{r}
# Compile results
ESS5 <- effectiveSize(samples5)
out5 <- summary(samples5)$quantiles
rownames(out5)<-names

ESS5
```

### Plot Posterior Density Individually
```{r}
beta <- NULL
for(l in 1:n.chains){
beta <- rbind(beta,samples5[[l]])
}
colnames(beta) <- names
for(j in 1:8){
hist(beta[,j],xlab=expression(beta[j]),ylab="Posterior density",
breaks=100,main=names[j])
}
```

### SSVS
```{r}
library(knitr)
Inc_Prob <- apply(beta!=0,2,mean)
Q <- t(apply(beta,2,quantile,c(0.5,0.05,0.95)))
out <- cbind(Inc_Prob,Q)
kable(round(out,2))
```
Summary: All are included with posterior probability one and are
thus clearly important predictors of price. SO we decided to use all features

### DIC-WAIC
```{r}
# Compute DIC-WAIC
DIC_WAIC_cal(model5)
```
## Model 6 (Using features that have strong correlation with the target variable)
We use features that have three stars.
```{r}
price.lm <- lm(price ~., data = df)

summary(price.lm)
```
There are 7 variables that have a strong correlation with the variable price (indicated by 3 stars). These seven variables include squareMeters, hasYard, hasPool, floors, cityPartRange, isNewBuilt, and hasStormProtector.

# Split dependent and independent variable 
Using 7 variables that have a strong correlation with the variable price
```{r}
X <- df[, c("squareMeters", "hasYard", "hasPool", "floors", "cityPartRange", "isNewBuilt", "hasStormProtector")]
Y <- df$price
names <- c("price", "Square Meters", "Has Yard", "Has Pool", "Floors", "City Part Range", "Is New Built", "Has Storm Protector")
```

# Delete Missing Value
```{r}
junk <- is.na(rowSums(X))
Y <- Y[!junk]
X <- X[!junk,]
```

# Standardize Covariates
```{r}
X <- as.matrix(scale(X))
```

## JAGS
# Put Data in JAGS Format
```{r}
n <- length(Y)
p <- ncol(X)

data <- list(Y=Y,X=X,n=n,p=p)
params <- c("alpha","beta")
burn <- 500
n.iter <- 2000
n.chains <- 3
```

# Make Jags Model
```{r}
model_string <- textConnection("model{
  # Likelihood
  for(i in 1:n){
    Y[i] ~ dnorm(alpha+mu[i],taue)
    mu[i] <- inprod(X[i,],beta[])
  }
  
  # Priors
  for(j in 1:p){
    beta[j] ~ dnorm(0,0.001)
  }
  
  alpha ~ dnorm(0,0.001)
  taue ~ dgamma(0.1, 0.1)
  
  # WAIC calculations
  for(i in 1:n){
    like[i] <- dnorm(Y[i],mu[i],taue)
  }

}")
```

# Set Initial Value
```{r}
inits = list()

inits$alpha = rnorm(1)
for(i in 1:p) {
    inits$beta[i] = rnorm(1)
}
inits$taue = 10
```

# Compile Model
```{r}
model6 <- jags.model(model_string, data = data, n.chains=n.chains, quiet=TRUE, inits = inits)
```

# Update Model
```{r}
update(model6, burn)
```

# Get Posterior Samples from the model
```{r}
samples6 <- coda.samples(model3, variable.names=params, n.iter=n.iter)
```

```{r}
DIC_WAIC_cal(model3)
```

