---
title: "Random Effects Code"
author: "Wilbert"
date: "2023-12-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load library
```{r}
library(coda)
library(rjags)
library(knitr)
library(purrr)
library(tidyr)
library(ggplot2)
set.seed(123)
```

# Load Data
```{r}
df <- read.csv('./ParisHousing.csv')
head(df)
```

## Model 1
# Set Dataframe as matrix and split dependent and independent variable 
```{r}
price <- as.matrix(df$price)
Y <- price
X <- cbind(df$squareMeters, df$numberOfRooms, df$hasYard, df$hasPool, df$floors, df$cityCode, df$numPrevOwners, df$made, df$isNewBuilt, df$hasStormProtector, df$basement, df$attic, df$garage, df$hasStorageRoom, df$hasGuestRoom, df$cityPartRange)
names <- c("price","Square Meters", "Number of Rooms", "Has Yard", "Has Pool", "Number of Floors", "City Code", "Number of Previous Owners", "Year Made", "is Newly Built", "Has Storm Protector", "Basement Area", "Attic Area", "Garage Area", "Has Storage Room", "Has Guest Room", "City Part Range")
```

# Delete Missing Value
```{r}
junk <- is.na(rowSums(X))
Y <- Y[!junk]
X <- X[!junk,]
```

# Standardize Covariates
```{r}
X <- as.matrix(scale(X))
```

## JAGS
# Put Data in JAGS Format
```{r}
n <- length(Y)
p <- ncol(X)

data <- list(Y=Y,X=X,n=n,p=p)
params <- c("alpha","beta")
burn <- 500
n.iter <- 5000
n.chains <- 3
thin <- 5
```

# Make Jags Model
```{r}
model_string <- textConnection("model{
  # Likelihood
  for(i in 1:n){
    Y[i] ~ dnorm(alpha+mu[i],taue)
    mu[i] <- theta[i] + inprod(X[i,],beta[])
  }
  
  # Random Effects 
  for(j in 1:n){
    theta[j] ~ ddexp(0,taue)
  }
  
  # Priors
  for(j in 1:p){
    beta[j] ~ dnorm(0,0.001)
  }
  
  alpha ~ dnorm(0,0.001)
  taue ~ dgamma(0.1, 0.1)
  
  # WAIC calculations
  for(i in 1:n){
    like[i] <- dnorm(Y[i],mu[i],taue)
  }

}")
```

# Set Initial Value
```{r}
inits = list()

inits$alpha = rnorm(1)
for(i in 1:p) {
    inits$beta[i] = rnorm(1)
}
for(i in 1:n) {
    inits$theta[i] = rnorm(1)
}
inits$taue = 10
```

# Compile Model
```{r}
model <- jags.model(model_string, data = data, n.chains=n.chains, quiet=TRUE, inits = inits)
```

# Update Model
```{r}
update(model, burn)
```

# Get Posterior Samples from the model
```{r}
samples <- coda.samples(model, variable.names=params, n.iter=n.iter, thin=thin)
```

# Plot MCMC Chain Trace and Features Posterior Density
```{r}
par(mar=c(1,1,1,1))
plot(samples)
```

# Check Convergence
```{r}
gelman.diag(samples)
```

# Show Alpha and Betas to the corresponding Features
```{r}
sum <- summary(samples)
rownames(sum$statistics) <- names
rownames(sum$quantiles) <- names
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

# Compile results
```{r}
ESS2 <- effectiveSize(samples)
out2 <- summary(samples)$quantiles
rownames(out2)<-names

ESS2
```

# Compute DIC & WAIC
```{r}
# DIC
dic <- dic.samples(model,n.iter=n.iter)

# WAIC
waic <- coda.samples(model, variable.names=c("like"), n.iter=n.iter)
like <- waic[[1]]
fbar <- colMeans(like)
P <- sum(apply(log(like),2,var))
WAIC <- -2*sum(log(fbar))+2*P
```

```{r}
dic
```
```{r}
WAIC
```
