---
title: "Latihan-UAS-1-Regression"
author: "Wilbert"
date: "2023-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data Description

https://www.statsmodels.org/dev/datasets/generated/longley.html

    1. TOTEMP - Total Employment
    2. GNPDEFL - GNP deflator
    3. GNP - GNP
    4. UNEMP - Number of unemployed
    5. ARMED - Size of armed forces
    6. POP - Population
    7. YEAR - Year (1947 - 1962)
    
# 2. Load libraries & data
```{r}
library(tidyverse)
library(rpart)
library(rpart.plot)
library(MLmetrics)
```


```{r}
data(longley)

head(longley)
```

# 3. Statistics Summary
```{r}
basicSummary <- function(df, dgts = 3) {
  m <- ncol(df) # column
  varNames <- colnames(df)
  varType <- vector("character", m)
  topLevel <- vector("character", m)
  topCount <- vector("numeric", m)
  missCount <- vector("numeric", m)
  levels <- vector("numeric", m)
  
  for (i in 1:m) {
    x <- df[,i]
    varType[i] <- class(x)
    xtab <- table(x, useNA = "ifany")
    levels[i] <- length(xtab)
    nums <- as.numeric(xtab)
    maxnum <- max(nums)
    topCount[i] <- maxnum
    maxIndex <- which.max(nums)
    lvls <- names(xtab)
    topLevel[i] <- lvls[maxIndex]
    missIndex <- which((is.na(x)) | (x == "") | (x == " "))
    missCount[i] <- length(missIndex)
  }
  n <- nrow(df) # Rows
  topFrac <- round(topCount/n, digits = dgts)
  missFrac <- round(missCount/n, digits = dgts)
  summaryFrame <- data.frame(variable = varNames, 
                             type = varType, 
                             levels = levels,
                             topLevel = topLevel,
                             topCount = topCount,
                             topFrac = topFrac,
                             missFreq = missCount,
                             missFrac = missFrac)
  return(summaryFrame)
}
```
```{r}
basicSummary(longley)
```
GNP deflator : numeric variable with 16 objects, with each object appearing once adn there's no missin value.
GNP : numeric variable with 16 objects, with each object appearing once adn there's no missin value.
Unemployed : numeric variable with 16 objects, with each object appearing once adn there's no missin value.
Armed Forces : numeric variable with 16 objects, with each object appearing once adn there's no missin value.
Population : numeric variable with 16 objects, with each object appearing once adn there's no missin value.
Employed : numeric variable with 16 objects, with each object appearing once adn there's no missin value.
Year : Integer variable with 16 objects, with each object appearing once adn there's no missin value.-> Should be treated as a Category

```{r}
summary(longley)
```
GNP deflator : numeric variable with Min of 83, Median of 100.60, Mean of 101.68, Max of 116.90, Q1 of 94.53, and Q3 of 111.25
GNP : numeric variable with Min of 234.3, Median of 231.4, Mean of 387.7, Max of 554.9, Q1 of 317.9, and Q3 of 454.1
Unemployed : numeric variable with Min of 187, Median of 314.4, Mean of 319.3, Max of 480.6, Q1 of 234.8, and Q3 of 384.2
Armed Forces : numeric variable with Min of 145.6, Median of 217.8, Mean of 260.7, Max of 359.4, Q1 of 229.8, and Q3 of 306.1
Population : numeric variable with Min of 107.6, Median of 116.8, Mean of 117.4, Max of 130.1, Q1 of 111.8, and Q3 of 122.3
Year : Integer variable with Min of 1947, Median of 1954, Mean of 1954, Max of 1962, Q1 of 1951, and Q3 of 1958
Employed : numeric variable with Min of 60.17, Median of 65.5, Mean of 65.32, Max of 70.55, Q1 of 62.71, and Q3 of 68.29

# 4. Data Preparation
- Change Year Variable to Numerical
- Create a subset for the numerical data

```{r}
longley$Year <- as.numeric(longley$Year)
```
```{r}
basicSummary(longley)
head(longley)
```

# 5. Check For Normality
```{r}
par(mfrow = c(2, 3))
hist(longley$GNP.deflator)
hist(longley$GNP)
hist(longley$Unemployed)
hist(longley$Armed.Forces)
hist(longley$Population)
hist(longley$Employed)
```
From the Graph above we can see that only the Employed Variable is not Normaly distributed thus will be excluded from the model

# 6. Split Data
```{r}
longley <- subset(longley, select = -c(Employed))
set.seed(123)
n <- nrow(longley)
train <- sample(n, round(0.7 * n))
longleyTrain <- longley[train,]
longleyTest <- longley[-train,]
```

# 7. Fit the Model
## 7A. Linear Regression Model
```{r}
LRModel1 <- lm(GNP ~ ., longleyTrain)
```
```{r}
res <- resid(LRModel1)
plot(fitted(LRModel1), res)
abline(0,0)
plot(density(res))
```

From the Residual and Density polt Above we can conclude that the LR1Model is Normally Distributed
```{r}
summary(LRModel1)
```
We will use top the variable with 1 star and above for the 2nd Linear Model
```{r}
LRModel2 <- lm(GNP ~ Unemployed + GNP.deflator + Year + Population, longleyTrain)
```

## 7B. Decission Tree Model
```{r}
DTModel1 <- rpart(GNP ~ ., longleyTrain)
DTModel1$variable.importance
```
Based on the Result ABove we can say that Decission Tree Model is not suitable

# 8. Model Validation
```{r}
predLR1 <- predict(LRModel1, longleyTest)
predLR2 <- predict(LRModel2, longleyTest)
```

# 9. Model Performance
```{r}
# Calculating RMSE
valLR1 <- RMSE(predLR1, longleyTest$GNP)
valLR2 <- RMSE(predLR2, longleyTest$GNP)

sprintf("RMSE of Linear Model 1 is %s", valLR1)
sprintf("RMSE of Linear Model 2 is %s", valLR2)

# Calculationg R2_Score
valLR1 <- R2_Score(predLR1, longleyTest$GNP)
valLR2 <- R2_Score(predLR2, longleyTest$GNP)

sprintf("R2_Score of Linear Model 1 is %s", valLR1)
sprintf("R2_Score of Linear Model 2 is %s", valLR2)
```
We can conclude that the best model is the Linear Regression Model 1 because it has the least RMSE and the highest R2_Score
