---
title: "New-Notes"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##  Exploring New Dataset

### Basic Sequence
1. Assess the general characteristics of the dataset. Ex :
    a. How many rows/observation/records we have?
    b. How many column/variables/features we have?
    c. What is the variables names and are they meaningfull? (meaningfull = menggabarkan isinya, jika tidak then ganti namanya)
    d. What is the type of each data. (numeric(quantitative) / categorical : (nominal(no ranks) / ordinal(have ranks)) (qualitative) / logical(boolean))
    e. How many unique values does each variable have? (value in a variable without counting for frequency or double) Ex : in age 10-12 have {10, 11, 12} as unique values.
    f. What is the value is the mode and what is it's frequency?
    g. Are there missing observation? if there are, then what is it's frequency?
    
2. Examine the descriptive statistics of each variable :
    - mean
    - median
    - mode
    - standart deviation
    - varians
    - range
    - quartile
    - min/max
    - IQR
    - Kurtosis
    - Skewness
    
    Descriptive Statistics is divided by 2 :
     a. Central Tendencies :
          - mean
          - median
          - mode
     b. Variability :
          - most use standart deviation
          
3. When possible make the visualization for the variable of interest (variable yang interesting) (maybe subjective) :
    a. can be from the data description
    b. can be from personal experience
    
4. When possible seatch for the anomally/outlier :
    a. explain the anomally/outlier (what and why)
    b. delete if make sense
    c. doesn't have to be deleted (if not needed)
    d. more in assignment
    
5. Look for the relation between variable of interest/key variable :
    a. don't just explain the relation but also the key variable first
    
6. Summarize data in the form of data dictionary 

## Data (Hands on)
```{r}
library(carData)
data("Chile")

```

search fot the dimension and the type of the variables
```{r}
dim(Chile)
str(Chile)
mean(Chile$population)
quantile(Chile$population, c(0.25, 0.5, 0.75))
summary(Chile)
```

Analysis result : the dataframe have 2700 objects and 8 variables...

How many Unique values a variable have
```{r}
# Unique
unique(Chile$region) # Return Factor

# Levels
levels(Chile$region) # Return Character

# Table
table(Chile$region) # Gives the Frequency too
```

User Defined Function to give basic summary
```{r}
basicSummary <- function(df, dgts) {
  m <- ncol(df) # column
  varNames <- colnames(df)
  varType <- vector("character", m)
  topLevel <- vector("character", m)
  topCount <- vector("numeric", m)
  missCount <- vector("numeric", m)
  levels <- vector("numeric", m)
  
  for (i in 1:m) {
    x <- df[,i]
    varType[i] <- class(x)
    xtab <- table(x, useNA = "ifany")
    levels[i] <- length(xtab)
    nums <- as.numeric(xtab)
    maxnum <- max(nums)
    topCount[i] <- maxnum
    maxIndex <- which.max(nums)
    lvls <- names(xtab)
    topLevel[i] <- lvls[maxIndex]
    missIndex <- which((is.na(x)) | (x == "") | (x == " "))
    missCount[i] <- length(missIndex)
  }
  n <- nrow(df) # Rows
  topFrac <- round(topCount/n, digits = dgts)
  missFrac <- round(missCount/n, digits = dgts)
  summaryFrame <- data.frame(variable = varNames, 
                             type = varType, 
                             levels = levels,
                             topLevel = topLevel,
                             topCount = topCount,
                             topFrac = topFrac,
                             missFraq = missCount,
                             missFrac = missFrac)
  return(summaryFrame)
}
```

```{r}
basicSummary(Chile)
```





