---
title: "Working with Text Data"
author: "Wilbert"
date: "2023-05-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fundamentals of text data analysis
Basic Steps : 
- Get the text to analyze
- Normalize the text by eliminating irrelevant details, cth :
    a. Remobe case : all text become upper or lower case.
    b. Remove useless character : punctuaion, symbols, numbers.
    c. Parse the text into words
    d. Remove non-informative stopwords
    e. Stem the text, removing word endings like "ing", "s", or "ed".
- Convert the normalized text to numbers 
- Apply mathematical analysis to the numbers
- Intepret the analysis results

Text analysis can be done in R using R native function or using tm or quanteda packages

Important base function : 
- nchar   : character count
- grep    : finds vector
- gsub    : replace substring
- tolower : to lower

# Creating Word Cloud

## Load Packages
```{r}
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
```

## Making word cloud

### Reading Text File

- 1st Way
```{r}
text <- readLines(file.choose())
```
- 2nd Way
```{r}
filePath <- "text-mining.txt"
text <- readLines(filePath)
```
### Looking data as corpus and inspecting
```{r}
docs <- Corpus(VectorSource(text))

inspect(docs)
```
### Text Transformation

```{r}
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")
```

### Text Cleaning
```{r}
# Convert the text to lower case
docs <- tm_map(docs, content_transformer(tolower))

# Remove numbers
docs <- tm_map(docs, removeNumbers)

# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))

# Remove your own stop word

# specify your stopwords as a character vector
docs <- tm_map(docs, removeWords, c("blabla1", "blabla2")) 

# Remove punctuations
docs <- tm_map(docs, removePunctuation)

# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)

# Text stemming
docs <- tm_map(docs, stemDocument)
```

### Build a term document matrix
```{r}
dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
d
```
### Generate Word Cloud
```{r}
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 4,
          max.words=200, random.order=FALSE, rot.per=0.20, 
          colors=brewer.pal(8, "Dark2"))
```





