---
title: "Session1_EDA"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Import Library and Data
```{r}
library(carData)

data(Davis)
ls()
Davis
```

User Defined Function to give basic summary
```{r}
basicSummary <- function(df, dgts) {
  m <- ncol(df) # column
  varNames <- colnames(df)
  varType <- vector("character", m)
  topLevel <- vector("character", m)
  topCount <- vector("numeric", m)
  missCount <- vector("numeric", m)
  levels <- vector("numeric", m)
  
  for (i in 1:m) {
    x <- df[,i]
    varType[i] <- class(x)
    xtab <- table(x, useNA = "ifany")
    levels[i] <- length(xtab)
    nums <- as.numeric(xtab)
    maxnum <- max(nums)
    topCount[i] <- maxnum
    maxIndex <- which.max(nums)
    lvls <- names(xtab)
    topLevel[i] <- lvls[maxIndex]
    missIndex <- which((is.na(x)) | (x == "") | (x == " "))
    missCount[i] <- length(missIndex)
  }
  n <- nrow(df) # Rows
  topFrac <- round(topCount/n, digits = dgts)
  missFrac <- round(missCount/n, digits = dgts)
  summaryFrame <- data.frame(variable = varNames, 
                             type = varType, 
                             levels = levels,
                             topLevel = topLevel,
                             topCount = topCount,
                             topFrac = topFrac,
                             missFraq = missCount,
                             missFrac = missFrac)
  return(summaryFrame)
}
```

```{r}
basicSummary(Davis, 3)
```

The Davis Data frame is 2 dimention with 200 objects and 5 variables with the description of
- sex : Categorical data with 2 unique values with "F" being the most often calue occuring 112 times. With no missing values
- weight : Numerical data with 54 unique values with 55 being the most often value occuring 9 times. With no missing values
- height : Numerical data with 43 unique values with 178 being the most often value occuring 12 times. With no missing values.
- repwt : Numerical data with 54 unique values with NA being the most often value occuring 17 times. With 17 missing values
- repht : Numerical data with 33 unique values with 165 being the most often value occuring 23 times. With 17 missing values


```{r}
summary(Davis)
```
The Descriptive Statistics of the Davis Datasets :

weight : Have minimum value of 39, median of 63, mean of 65, max of 166, the 1st Quartile is 55, and the 3rt Quartile is 74 
height : Have minimum value of 57, median of 169.5, mean of 170, max of 197, the 1st Quartile is 164, and the 3rt Quartile is 177.2 
repwt : Have minimum value of 41, median of 63, mean of 65.62, max of 124, the 1st Quartile is 55, and the 3rt Quartile is 73 and missing values of 17.
repht : Have minimum value of 148, median of 168, mean of 168.5, max of 200, the 1st Quartile is 160.5, and the 3rt Quartile is 175 and missing value of 17.

```{r}
plot(Davis, las = 1)
```
```{r}
par(mfrow = c(2,2))

x <- Davis$height
y <- Davis$weight

plot(x, y, xlab = "Height", ylab = "Weight", ylim = c(40, 170))
index <- which(y > 140)
points(x[-index], y[-index], pch = 16)
points(x[index], y[index], pch = 18, cex = 2, col = "#009999")
title("Plot Weight vs Height")

plot(Davis$height, Davis$repht, xlab = "Measured Height in cm", ylab = "Reported Height in cm", xlim = c(50, 200), col = "#0000FF", main = "Plot Measured Height vs Reported Height")

plot(Davis$weight, Davis$repwt, xlab = "Measured Weight in kg", ylab = "Reported Weight in kg", ylim = c(40, 130), col = "red", main = "Plot Measured Weight vs Reported Weight")

plot(Davis$repwt, Davis$repht, xlab = "Reported Weight in kg", ylab = "Reported Height in cm", ylim = c(140, 200), col = "green", main = "Plot Reported Weight vs Reported Height")
```
Weight vs Height Description :
  from the plot above we can see that the heavier someone is then they are slightly (positive corellation)




